Towards caring 15-minute neighbourhoods

### 2024-05-27 data/02-preliminary-analysis.qmd
part 2 removed weighted calculations in 'preliminary-analysis'

test_all <- ttm_care_walk_15 |>
  sp_avail_detailed(o_id = from_id,
                    d_id = to_id, 
                    pop = pop_dwell, #pop assumed at each res. parcel -- it is the pop in the DA divided by the number of dwellings in the DA. 
                    opp = f, 
                    r = f, f = f, alpha = 1) |> 
  group_by(from_id, Care_Category, Care_Category_Specific) |> summarize(V_isc_w = sum(V_ij, na.rm = TRUE))

access_15 <- access_15 |> left_join(test_all, by=c("from_id", "Care_Category", "Care_Category_Specific")) |> ungroup()
rm(test_all)
access_15$V_isc_w |> sum(na.rm = TRUE) #sums to 2206 opportunities -- out of 2225 total opportunities. some opportunities canot be reached wtihin 15 min walk! 
```

@@

           V_isc_w = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic")))) |>
  mutate(from_id = as.character(from_id))

@@

access_15 <- bind_rows(access_15_zeros, access_15) |>
  group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(GeoUID = first(GeoUID),
            copp_isc = sum(copp_isc),
            V_isc_w = sum(V_isc_w))
rm(access_15_zeros)
#check
access_15$V_isc_w |> sum(na.rm=T)
```

@@

```{r}
# by care category, for parcels 
access_15_bycategory <- access_15 |> group_by(from_id, Care_Category)  |> 
  summarize(GeoUID = first(GeoUID),
            copp_ic = sum(copp_isc, na.rm=T), #the number of opportunities that can be reached by category from each parcel
            copp_ic_w = sum(copp_isc, na.rm=T), #the number of opportunities that can be reached by category from each parcel
            V_ic_w = sum(V_isc_w, na.rm = T))
#check
access_15_bycategory$V_ic_w |> sum(na.rm=T)
```

```{r}
#by DA and mean:
access_15_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |> 
  group_by(GeoUID) |>
  summarise(copp_iscDA_med = median(copp_isc), #the median number of opportunities (all) that can be reached from each parcel in a DA
            copp_iscDA_sd = sd(copp_isc), #the SD
    copp_iscDA_w_med = median(copp_isc_w), #the median number of opportunities (all) that can be reached from each parcel in a DA
            copp_iscDA_w_sd = sd(copp_isc_w), #the SD
            V_w_iscDA = sum(V_isc_w), #the number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            vv_w_iscDA = sum(V_isc_w)/first(Population), #the per-capita number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_)) |>
  tidyr::drop_na(Population)

@@

# now by category
access_15_bycategory_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population","Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |>
  group_by(GeoUID, Care_Category) |>
  summarise(copp_ic_med = median(copp_isc), #the median number of opportunities that can be reached by category from each parcel in a DA
            copp_ic_sd = sd(copp_isc), #the SD
            copp_ic_w_med = median(copp_isc_w), #the median number of opportunities that can be reached by category from each parcel in a DA
            copp_ic_w_sd = sd(copp_isc_w), #the SD
            V_ic_w = sum(V_isc_w), #the number of total spatially available opportunities that can be reached by category from each parcel in a DA
            vv_ic_w = sum(V_isc_w)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by category from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_) ) |>
  tidyr::drop_na(Population)

#now by specific category
access_15_byspecificcategory_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population","Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |>
  group_by(GeoUID, Care_Category, Care_Category_Specific) |>
  summarise(copp_isc_med = median(copp_isc), #the median number of opportunities that can be reached by specific category from each parcel in a DA
            copp_isc_sd = sd(copp_isc), #the SD
            copp_isc_w_med = median(copp_isc_w), #the median number of opportunities that can be reached by specific category from each parcel in a DA
            copp_isc_w_sd = sd(copp_isc_w), #the SD
            V_isc_w = sum(V_isc_w), #the number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
            vv_isc = sum(V_isc_w)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_) ) |>
  tidyr::drop_na(Population)

#checks
# access_15_DA |> summary()
access_15_bycategory_DA |> summary()
access_15_DA$V_w_iscDA |> sum(na.rm=T)
access_15_bycategory_DA$V_ic_w |> sum(na.rm=T)
access_15_byspecificcategory_DA$V_isc_w |> sum(na.rm=T)

#RECALL, categories are either 
# - "Dependent-centric" and include "Community or Recreation Centre", "Daycare or EarlyON", "Long-Term Care or Retirement Home", "Park", "School", "Senior Centre"
# - "Errand-centric" and include: "Bank or ATM", "Library", "Post Office"
# - "Grocery-centric" and include: "Convenience Store", "Grocer"
# - "Health-centric" and include: "Dentist", "Hospital or Clinic", "Pharmacy"
```
```{r}
access_15_byspecificcategory_DA |> group_by(Care_Category_Specific) |> summarise(V_isc_w = sum(V_isc_w))
care_dest |> st_drop_geometry() |> group_by(Care_Category_Specific) |> 
  summarise(Dest_count_weighted = sum(Dest_count_weighted),
            count = n())

@@

```{r}
#calculating entropy, here I do it by the 14 specific categories. This is a way to consider their diversity within and between categories. Because they have been weighted before calc. the cumulative opp counts, we can proceed without anymore adjustments.
step1 <- access_15_byspecificcategory_DA  |> group_by(GeoUID) |> 
  summarise(OPPi = sum(copp_isc_med, na.rm=T) + sum(copp_isc_sd, na.rm=T),
            OPPi_w = sum(copp_isc_w_med, na.rm=T) + sum(copp_isc_w_sd, na.rm=T),
            Vi_w = sum(V_isc_w))

@@

step2 <- step2 |> mutate(top1_copp = (copp_isc_med+copp_isc_sd) / OPPi * log((copp_isc_med+copp_isc_sd) / OPPi),
             top1_copp = ifelse(is.nan(top1_copp), 0, top1_copp),
             top1_coppw = (copp_isc_w_med+copp_isc_w_sd) / OPPi_w * log((copp_isc_w_med+copp_isc_w_sd) / OPPi_w),
             top1_coppw = ifelse(is.nan(top1_coppw), 0, top1_coppw),
             top1_Vw= (V_isc_w) / Vi_w * log((V_isc_w) / Vi_w),
             top1_Vw = ifelse(is.nan(top1_Vw), 0, top1_Vw),)  |> 
  group_by(GeoUID) |>
  summarise(entropy_copp_isc_medsd = sum(top1_copp) / -log(14),
            entropy_copp_isc_medsd_w = sum(top1_coppw) / -log(14),
            entropy_V_isc_w = sum(top1_Vw) / -log(14))

#join this to access_15_bycategory_DA
access_15_bycategory_DA <- access_15_bycategory_DA |> left_join(step2, by="GeoUID")
```
COMEBACKHERE
```{r}
#the number of opportunities from each categories
all_cats <- access_15_bycategory_DA |> 
  mutate(copp_all = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T),
         copp_all_w = sum(copp_ic_w_med, na.rm=T) + sum(copp_ic_w_sd, na.rm=T)) |>
  group_by(GeoUID) |>
         summarise(summed_copp_all = sum(copp_all),
                   summed_copp_all_w = sum(copp_all_w),
                   summed_Vic_w = sum(V_ic_w, na.rm=T),
                   Population = mean(Population),
                   Care_Access_Status1 = case_when(summed_copp_all < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all >= 4 & summed_copp_all < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all >= 25 ~ "[25+ des.) Very caring"),
                   Care_Access_Status1_w = case_when(summed_copp_all_w < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all_w >= 4 & summed_copp_all_w < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all_w >= 25 ~ "[25+ des.) Very caring"),

                   Care_Access_Status2 = case_when(all(copp_all >= 2 & copp_all < 4) ~ "- complete (2 to 4 options per cat.)", 
                                                   all(copp_all >= 4) ~ "- very complete (4 or more options per cat.)"),
                   Care_Access_Status2_w = case_when(all(copp_all_w >= 2 & copp_all_w < 4) ~ "- complete (2 to 4 options per cat.)", 
                                                   all(copp_all_w >= 4) ~ "- very complete (4 or more options per cat.)"),
                   entropy_copp_isc_medsd = first(entropy_copp_isc_medsd),
                   entropy_copp_isc_medsd_w = first(entropy_copp_isc_medsd_w),
                   entropy_V_isc_w = first(entropy_V_isc_w)) |>
  mutate(Care_Access_Status2 = ifelse(is.na(Care_Access_Status2), "", Care_Access_Status2),
         Care_Access_Status2_w = ifelse(is.na(Care_Access_Status2_w), "", Care_Access_Status2_w),
         Care_Access_Status = paste(Care_Access_Status1,Care_Access_Status2, sep=" "),
         Care_Access_Status_w = paste(Care_Access_Status1_w,Care_Access_Status2_w, sep=" "),
         vvi = summed_Vic_w/Population)

all_cats$summed_Vic_w |> sum() #only 2206  opps can be reached out of the total 2225, so that's what the numbers are based on.
care_dest|> nrow()

@@

tmap_save(all_cats_copp_plot, file="all_cats_copp_plot.tiff")

all_cats_copp_weighted_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("Care_Access_Status_w",
              border.alpha = 0.2,
              palette = c("darkred", "orangered", "orange", "yellow", "forestgreen", "lightgreen"),
              title = "Number of care destinations (all - weighted) spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12
            # legend.title.size = 0.7,
            # legend.text.size = 0.6
            )
tmap_save(all_cats_copp_weighted_plot, file="all_cats_copp_weighted_plot.tiff")
```

@@

tmap_save(all_cats_entropy_copp_plot, file="all_cats_entropy_copp_plot.tiff")

all_cats_entropy_copp_weighted_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("entropy_copp_isc_medsd_w",
              border.alpha = 0.2,
              palette = "Reds",
              title = "Diversity of (all - weighted) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_entropy_copp_weighted_plot, file="all_cats_entropy_copp_weighted_plot.tiff")
```

@@

```{r}
all_cats |> group_by(Care_Access_Status_w) |> summarise(min = min(summed_copp_all_w),
                                                        median = median(summed_copp_all_w),
                                                        max = max(summed_copp_all_w))
all_cats |> group_by(Care_Access_Status) |> summarise(min = min(summed_copp_all),
                                                        median = median(summed_copp_all),
                                                        max = max(summed_copp_all))
all_cats$entropy_copp_isc_medsd|> summary()
all_cats$entropy_copp_isc_medsd_w |> summary()
all_cats$Care_Access_Status_w |> unique()
```

```{r}
#adding entropy labels
all_cats <- all_cats |> 
  mutate(bivariate_entropy_w = 
           case_when( ((Care_Access_Status_w == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status_w =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status_w =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                         entropy_copp_isc_medsd_w >= 0.7154) ~ "Caring, complete and amenity diverse",  #the median diversity,

                     ((Care_Access_Status_w == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status_w =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status_w =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                        entropy_copp_isc_medsd_w < 0.7154) ~ "Caring, complete but not amenity diverse",

                     ((Care_Access_Status_w == "[0-04 des.) Not caring " |
                       Care_Access_Status_w == "[04-25 des.) Caring ")  &
                        entropy_copp_isc_medsd_w >= 0.7154) ~ "Not caring or not complete, but amenity diverse",

                     ((Care_Access_Status_w == "[0-04 des.) Not caring "  |
                       Care_Access_Status_w == "[04-25 des.) Caring ")  &
                        entropy_copp_isc_medsd_w < 0.7154) ~ "Not caring, complete, or amenity diverse"),

         bivariate_entropy = 
           case_when( ((Care_Access_Status == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                         entropy_copp_isc_medsd >= 0.7418) ~ "Caring, complete and amenity diverse",  #the median diversity,

@@

```{r}
all_cats_bivariate_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("bivariate_entropy_w",
              border.alpha = 0.2,
              palette = c("yellowgreen","yellow","blue", "#FFFFDF"),
              title = "Diversity of (all - weighted) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

@@

```{r}
all_cats_entropy_copp_weighted_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("entropy_copp_isc_medsd_w",
              border.alpha = 0.2,
              palette = "Reds", style="quantile", n=4,
              title = "Diversity of (all - weighted) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_entropy_copp_weighted_plot, file="all_cats_entropy_copp_weighted_plot.tiff")
```

@@

Questions:

- what is a reasonable travel time? Is 15 minute walk indeed fine for our purpose?

- is 1 potential opportunity enough per subcategory?

- Should some subcategories be substitutes? E.g., 1 daycare OR EARLYON centre is enough, not 1 of each.

- Are there ways the categories can overlap to tell a story? Maybe to include google maps images. eg. 
  - communities where children+errand+grocery centric care is easier -- especially for women
  - health_centric+Elder_centric -- people with higher health care needs

- care trips 30% of all trips made. they are essential and have

>>>>

test_all <- ttm_care_walk_15 |>
  sp_avail_detailed(o_id = from_id,
                    d_id = to_id, 
                    pop = pop_dwell, #pop assumed at each res. parcel -- it is the pop in the DA divided by the number of dwellings in the DA. 
                    opp = f, 
                    r = f, f = f, alpha = 1) |> 
  group_by(from_id, Care_Category, Care_Category_Specific) |> summarize(V_isc = sum(V_ij, na.rm = TRUE))

access_15 <- access_15 |> left_join(test_all, by=c("from_id", "Care_Category", "Care_Category_Specific")) |> ungroup()
rm(test_all)
access_15$V_isc |> sum(na.rm = TRUE) #sums to 2206 opportunities -- out of 2225 total opportunities. some opportunities canot be reached wtihin 15 min walk! 
```

@@

           V_isc = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic")))) |>
  mutate(from_id = as.character(from_id))

@@

access_15 <- bind_rows(access_15_zeros, access_15) |>
  group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(GeoUID = first(GeoUID),
            copp_isc = sum(copp_isc),
            V_isc = sum(V_isc))
rm(access_15_zeros)
#check
access_15$V_isc |> sum(na.rm=T)
```

@@

```{r}
# by care category, for parcels 
access_15_bycategory <- access_15 |> group_by(from_id, Care_Category)  |> 
  summarize(GeoUID = first(GeoUID),
            copp_ic = sum(copp_isc, na.rm=T), #the number of opportunities that can be reached by category from each parcel
            V_ic = sum(V_isc, na.rm = T))
#check
access_15_bycategory$V_ic |> sum(na.rm=T)
```

```{r}
#by DA and mean:
access_15_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |> 
  group_by(GeoUID) |>
  summarise(copp_iscDA_med = median(copp_isc), #the median number of opportunities (all) that can be reached from each parcel in a DA
            copp_iscDA_sd = sd(copp_isc), #the SD
    copp_iscDA_w_med = median(copp_isc_w), #the median number of opportunities (all) that can be reached from each parcel in a DA
            copp_iscDA_w_sd = sd(copp_isc_w), #the SD
            V_iscDA = sum(V_isc), #the number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            vv_iscDA = sum(V_isc)/first(Population), #the per-capita number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_)) |>
  tidyr::drop_na(Population)

@@

# now by category
access_15_bycategory_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population","Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |>
  group_by(GeoUID, Care_Category) |>
  summarise(copp_ic_med = median(copp_isc), #the median number of opportunities that can be reached by category from each parcel in a DA
            copp_ic_sd = sd(copp_isc), #the SD
            V_ic = sum(V_isc), #the number of total spatially available opportunities that can be reached by category from each parcel in a DA
            vV_ic = sum(V_isc)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by category from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_) ) |>
  tidyr::drop_na(Population)

#now by specific category
access_15_byspecificcategory_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population","Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |>
  group_by(GeoUID, Care_Category, Care_Category_Specific) |>
  summarise(copp_isc_med = median(copp_isc), #the median number of opportunities that can be reached by specific category from each parcel in a DA
            copp_isc_sd = sd(copp_isc), 
            V_isc = sum(V_isc), #the number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
            vv_isc = sum(V_isc)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_) ) |>
  tidyr::drop_na(Population)

#checks
# access_15_DA |> summary()
access_15_bycategory_DA |> summary()
access_15_DA$V_iscDA |> sum(na.rm=T)
access_15_bycategory_DA$V_ic |> sum(na.rm=T)
access_15_byspecificcategory_DA$V_isc |> sum(na.rm=T)

#RECALL, categories are either 
# - "Dependent-centric" and include "Community or Recreation Centre", "Daycare or EarlyON", "Long-Term Care or Retirement Home", "Park", "School", "Senior Centre"
# - "Errand-centric" and include: "Bank or ATM", "Library", "Post Office"
# - "Grocery-centric" and include: "Convenience Store", "Grocer"
# - "Health-centric" and include: "Dentist", "Hospital or Clinic", "Pharmacy"
```
```{r}
access_15_byspecificcategory_DA |> group_by(Care_Category_Specific) |> summarise(V_isc = sum(V_isc))
care_dest |> st_drop_geometry() |> group_by(Care_Category_Specific) |> 
  summarise(count = n())

@@

```{r}
#calculating entropy, here I do it by the 14 specific categories. This is a way to consider their diversity within and between categories. Because they have been weighted before calc. the cumulative opp counts, we can proceed without anymore adjustments.
step1 <- access_15_byspecificcategory_DA  |> group_by(GeoUID) |> 
  summarise(OPPi = sum(copp_isc_med, na.rm=T) + sum(copp_isc_sd, na.rm=T),
            Vi = sum(V_isc))

@@

step2 <- step2 |> mutate(top1_copp = (copp_isc_med+copp_isc_sd) / OPPi * log((copp_isc_med+copp_isc_sd) / OPPi),
             top1_copp = ifelse(is.nan(top1_copp), 0, top1_copp),
             top1_V= (V_isc) / Vi * log((V_isc) / Vi),
             top1_V = ifelse(is.nan(top1_V), 0, top1_V),)  |> 
  group_by(GeoUID) |>
  summarise(entropy_copp_isc_medsd = sum(top1_copp) / -log(14),
            entropy_V_isc = sum(top1_V) / -log(14))

#join this to access_15_bycategory_DA
access_15_bycategory_DA <- access_15_bycategory_DA |> left_join(step2, by="GeoUID")
```

```{r}
#the number of opportunities from each categories
all_cats <- access_15_bycategory_DA |> 
  mutate(copp_all = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T)) |>
  group_by(GeoUID) |>
         summarise(summed_copp_all = sum(copp_all),
                   summed_Vic = sum(V_ic, na.rm=T),
                   Population = mean(Population),
                   Care_Access_Status1 = case_when(summed_copp_all < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all >= 4 & summed_copp_all < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all >= 25 ~ "[25+ des.) Very caring"),
                   Care_Access_Status2 = case_when(all(copp_all >= 2 & copp_all < 4) ~ "- complete (2 to 4 options per cat.)", 
                                                   all(copp_all >= 4) ~ "- very complete (4 or more options per cat.)"),
                   entropy_copp_isc_medsd = first(entropy_copp_isc_medsd),
                   entropy_V_isc = first(entropy_V_isc)) |>
  mutate(Care_Access_Status2 = ifelse(is.na(Care_Access_Status2), "", Care_Access_Status2),
         Care_Access_Status = paste(Care_Access_Status1,Care_Access_Status2, sep=" "),
         vvi = summed_Vic/Population)

all_cats$summed_Vic |> sum() #only 2206  opps can be reached out of the total 2225, so that's what the numbers are based on.
care_dest|> nrow()

@@

tmap_save(all_cats_copp_plot, file="all_cats_copp_plot.tiff")
```

@@

tmap_save(all_cats_entropy_copp_plot, file="all_cats_entropy_copp_plot.tiff")
```

@@

```{r}
all_cats |> group_by(Care_Access_Status) |> summarise(min = min(summed_copp_all),
                                                        median = median(summed_copp_all),
                                                        max = max(summed_copp_all))
all_cats$entropy_copp_isc_medsd|> summary()
all_cats$Care_Access_Status |> unique()
```

```{r}
#adding entropy labels
all_cats <- all_cats |> 
  mutate(bivariate_entropy = 
           case_when( ((Care_Access_Status == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                         entropy_copp_isc_medsd >= 0.7418) ~ "Caring, complete and amenity diverse",  #the median diversity,

@@

```{r}
all_cats_bivariate_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("bivariate_entropy",
              border.alpha = 0.2,
              palette = c("yellowgreen","yellow","blue", "#FFFFDF"),
              title = "Diversity of (all) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

@@

@@

<!-- Questions: -->

<!-- - what is a reasonable travel time? Is 15 minute walk indeed fine for our purpose? -->

<!-- - is 1 potential opportunity enough per subcategory?  -->

<!-- - Should some subcategories be substitutes? E.g., 1 daycare OR EARLYON centre is enough, not 1 of each.  -->

<!-- - Are there ways the categories can overlap to tell a story? Maybe to include google maps images. eg.  -->
<!--   - communities where children+errand+grocery centric care is easier -- especially for women -->
<!--   - health_centric+Elder_centric -- people with higher health care needs -->

<!-- - care trips 30% of all trips made. they are essential and have -->

### 2024-05-27 paper/Manuscript.qmd
part 2 removed weighted calculations in 'preliminary-analysis'

@@

BONUS! Per capita spatial availability. This is not discussed in the methods. The spatial availability calculation assumes population is 'population density' and weighted opportunity count. With this map we can start thinking about per capita availability of opportunities. 
```{r plot-figure6, out.width=600, out.height=600}
#| label: fig-Fig6
#| fig-cap: "Per capita spatial availability of care opportunities. Uses weighted care opportunities, population density as competition, and 15 minute binary cut-off. Aggregates parcel level spatial availability at the DA level (by summation) and divides by the DA population. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_vvi_plot.tiff"))
```

`r word_comment("If you're interested in how the unweighted results look like for the same maps see the appendix. I've also calculated the cycling travel matrices as well as the 'no dependent' i.e., the same analysis but now only considering 3 care categories: health, errand, and grocery. For the time being though let's focus on this.", " ")`

@@

\pagebreak

# Appendix

```{r plot-figure7, out.width=600, out.height=600}
#| label: fig-Fig7
#| fig-cap: "UNWEIGHTED -- The cumulative opportunity measure. The number of care destinations that can be reached, per DA, within 15 min walk. Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection). Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_plot.tiff"))
```

```{r plot-figure8, out.width=600, out.height=600}
#| label: fig-Fig8
#| fig-cap: "UNWEIGHTED -- The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_plot.tiff"))
```

>>>>

Self organizing map (**SOM**) is used to profile the types of neighbourhoods by their magnitude in 15-minute walking accessibility to caring opportunities, diversity in these opportunities and socio-economic characteristics. SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space. It is a dimensionality reduction method as it applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we intuitively profile.

@@

@@

<!-- `r word_comment("If you're interested in how the unweighted results look like for the same maps see the appendix. I've also calculated the cycling travel matrices as well as the 'no dependent' i.e., the same analysis but now only considering 3 care categories: health, errand, and grocery. For the time being though let's focus on this.", " ")` -->
